---
title: "Peak calling"
author: "Shengtong Han"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## Region level peak calling 
Basic idea: identify regions in the whole genome which contains one peak ideally. Thus the regions should be neigher too large nor too short. Currently use 20 bp.  


### Step 1: Base level coverage calling
Start with .bam file. The fist step is to call 5' end  coverage at base level 
```{r base level coverage calling, echo=T, eval=F}
prefix1=Jump-positive
prefix2=mm9.umi_encoded_adaptor_removed_no_mismatch.sorted.dedup.bam
samtools view -b -F 20 -o $prefix1.$prefix2.plus  $prefix1.$prefix2
samtools view -b -f 16 -o $prefix1.$prefix2.minus  $prefix1.$prefix2
echo get minus and plus strand
############## sort strands ################
samtools sort $prefix1.$prefix2.plus $prefix1.$prefix2.plus.sorted
samtools sort $prefix1.$prefix2.minus $prefix1.$prefix2.minus.sorted
echo sort two strands
########### call coverage at base level ###############
bedtools genomecov -5 -d -ibam $prefix1.$prefix2.plus.sorted.bam  -g mm9.genome > $prefix1.$prefix2.plus.cove.base.level
echo get coverage for plus strand

bedtools genomecov -5 -d -ibam $prefix1.$prefix2.minus.sorted.bam  -g mm9.genome > $prefix1.$prefix2.minus.cove.base.level
echo get coverage for minus strand

```

### Setp 2: Split whole genome into small windows of equal size
### Step 3: Calculate p values for each window
### step 4: Summarize the results with Statistics of interest


## Plot jump distribution 
### Yuwen's pipleline via sos 

sos run ../code/170509_plot_coverage_for_bam_files_using_FDR_0.5_tab-seq_sites_test_RCC_han.sos  'get_coverage' --data_folder ../../project2/xinhe/Shengtong/Shengtong_JumpSeq/5hmC_48ng_downsample/  --file_name 48ng_11rep_merge.sorted.bam  test.bam --plus_reference_regions ../other_annotations/other_hmc_data_set/GSM882244_mESC.hmC_sites.FDR_0.0484.mm9.txt_minus.extend_100bp_separated.bed --minus_reference_regions ../other_annotations/other_hmc_data_set/GSM882244_mESC.hmC_sites.FDR_0.0484.mm9.txt_minus.extend_100bp_separated.bed --reference_sites_name GSM882244_FDR_0.0484  &

* run the code in the `processing_command` directory in `Tiao_Seq`
* bedtools must be version of v2.17.0

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
